ASAH1 rGBM Manuscript ELISA Figure

### 2022-12-15 results/elisa_plasma_proteomics/manuscript_elisa_figure_final.Rmd
[ADD] Kaplan-Meir plots and lm of time to recurrence

# Data I/O and Wrangling
installIfNeeded("data.table")
installIfNeeded("dplyr")
# Visualization
installIfNeeded("ggplot2")
installIfNeeded("grid")
installIfNeeded("gridExtra")
installIfNeeded("ggpubr")
# Table Visualization
installIfNeeded("gt")
# Statistics
installIfNeeded("rstatix")
installIfNeeded("broman")
# Machine Learning 
installIfNeeded("caret")
installIfNeeded("plotROC")
installIfNeeded("mixOmics", "BIO")
```

@@

<!-- # Distribution of ELISA Measurements -->
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=21, include=FALSE}
# iGBM vs rGBM ----
## Build gg dataframe for plotting ----
data_long %>%
  dplyr::filter(!is.na(manifestation_type) | patient_state==0) %>%
  dplyr::mutate(log2_conc = log2(conc+1)) %>%
  dplyr::group_by(gene, Number.x) %>%
  dplyr::mutate(Manifestation = plyr::mapvalues(manifestation_type, from=c("Initial manifestation", "Second manifestation"), to=c("iGBM", "rGBM"))) %>%
  dplyr::mutate(Manifestation = ifelse(is.na(Manifestation), "Ctrl", Manifestation)) %>%
  dplyr::mutate(Manifestation = factor(Manifestation, levels = c("Ctrl", "iGBM", "rGBM"))) -> pdata
pdata <- as.data.table(pdata)

@@

## Build the plot ----
ggpaired(data = pdata, id = "Number.x", x = "Manifestation", y = "log2_conc", fill = "Manifestation", palette = "jco", line.color = "grey", line.size = 0.4) +
  # stat_compare_means(paired = TRUE, label.x = 1.5, geom = "label") +
  # stat_compare_means(comparisons = my_comparisons, group.by="gene", label.y = c(4, 7, 7), label.x = c(1, 2, 3)) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata$Manifestation)), "Dark2"), labels = c("Ctrl", "Initial manifestation", "Second manifestation")) +
  # geom_jitter(aes(label=Number.x)) +
  # ggtitle("Initial vs Recurrence") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="GBM\nOccurence", ncol=1),
         col=guide_legend(title="Healthy Control Baseline")) -> p1

p1 + 
  stat_pvalue_manual(stat_test, label = "{p.signif} p = {p}", step.group.by = "gene", step.increase = 0.2, hide.ns = T, size=4, bracket.size=1) +
  geom_point(data=data.frame(Manifestation=c("iGBM", "iGBM", "iGBM", "iGBM"), gene=c("ASAH1", "SYNEMIN", "GPNMB", "MMP-9" ), y=c(8, 18, 9, 12)),
             aes(x=Manifestation, y=y, group=gene), alpha=0) -> p1

@@

@@

# Master figure ----
# subfigure_6 <- ggarrange(p1, p2, p3, ncol = 3)
subfigure_6 <- p1
subfigure_6
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a.png", plot = subfigure_6, width = 21, height = 11, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a.svg", plot = subfigure_6, width = 21, height = 11, dpi = 200)
```

@@

train_df_Sub$Number.x <- as.factor(train_df_Sub$Number.x)
train_df_Sub$Time.to.reccurence..days. <- as.numeric(train_df_Sub$Time.to.reccurence..days.)
lm_ttr <- lmerTest::lmer(Time.to.reccurence..days. ~ SYNEMIN + GPNMB + ASAH1 + MMP.9 + (1 | Type.of.resection), data=train_df_Sub )
summary(lm_ttr)
anova(lm_ttr)

lm_ttr <- lm(Time.to.reccurence..days. ~ SYNEMIN + GPNMB + ASAH1 + MMP.9, data=train_df_Sub )
summary(lm_ttr)
# car::Anova(lm_ttr, type="III")

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.png", plot = p_ratio_ttr, width = 15, height = 9, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.svg", plot = p_ratio_ttr, width = 15, height = 9, dpi = 200)

@@

## Only use proteins for prediction
data_wide %>%
  mutate(n = row_number()) %>%
  dplyr::select(Number.x, class, n, ASAH1, `MMP-9`, SYNEMIN, GPNMB, `Tumor localization`, Sex, `Age at surgery (years)`, `Type of resection`) -> data_wide

@@

@@

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c.png", plot = g, width = 11, height = 8.5, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c.svg", plot = g, width = 11, height = 8.5, dpi = 200)
subfigure_1 <- g
```

@@

>>>>

# Data I/O and Wrangling
installIfNeeded("data.table")
installIfNeeded("dplyr")
# Visualization
installIfNeeded("ggplot2")
installIfNeeded("grid")
installIfNeeded("gridExtra")
installIfNeeded("ggpubr")
# Table Visualization
installIfNeeded("gt")
# Statistics
installIfNeeded("rstatix")
installIfNeeded("broman")
installIfNeeded("pwr")
# Survival Analysis
installIfNeeded("survival")
installIfNeeded("survminer")
# Machine Learning 
installIfNeeded("caret")
installIfNeeded("plotROC")
installIfNeeded("mixOmics", "BIO")
```

@@

<!-- # Distribution of ELISA Measurements -->
```{r echo=FALSE, fig.height=11, fig.width=21, message=FALSE, warning=FALSE}
# iGBM vs rGBM ----
## Build gg dataframe for plotting ----
data_long %>%
  dplyr::filter(!is.na(manifestation_type) | patient_state==0) %>%
  dplyr::mutate(log2_conc = log2(conc+1)) %>%
  dplyr::group_by(gene, Number.x) %>%
  dplyr::mutate(Manifestation = plyr::mapvalues(manifestation_type, from=c("Initial manifestation", "Second manifestation"), to=c("iGBM", "rGBM"))) %>%
  dplyr::mutate(Manifestation = ifelse(is.na(Manifestation), "Ctrl", Manifestation)) %>%
  dplyr::mutate(Manifestation = factor(Manifestation, levels = c("Ctrl", "iGBM", "rGBM"))) -> pdata
pdata <- as.data.table(pdata)

@@

## Build the plot ----
colfunc <- colorRampPalette(c("red","yellow","springgreen","royalblue"))
ttr_categorizer <- function(data) {
  data <- as.numeric(data)
  if (is.na(data)){
    return(NA)
  }
  if (data >= 100 & data < 200){
    return("100-200")
  } else if (data >= 200 & data < 300){
    return("200-300")
  } else if (data >= 300 & data < 400){
    return("300-400")
  } else if (data >= 400 & data < 500){
    return("400-500")
  } else if (data >= 500 & data < 600){
    return("500-600")
  }
}
# ttr_categorizer <- function(data) {
#   data <- as.numeric(data)
#   if (is.na(data)){
#     return(NA)
#   }
#   if (data >= 100 & data < 150){
#     return("100-150")
#   } else if (data >= 150 & data < 200){
#     return("150-200")
#   } else if (data >= 200 & data < 250){
#     return("200-250")
#   } else if (data >= 250 & data < 300){
#     return("250-300")
#   } else if (data >= 300 & data < 350){
#     return("300-350")
#   } else if (data >= 350 & data < 400){
#     return("350-400")
#   } else if (data >= 400 & data < 450){
#     return("400-450")
#   } else if (data >= 450 & data < 500){
#     return("450-500")
#   } else if (data >= 500 & data < 550){
#     return("500-550")
#   } else if (data >= 550 & data < 600){
#     return("550-600")
#   }
# }
# ttr_categorizer <- function(data, split_value) {
#   if (data <= split_value) {
#     return(0)
#   } else {
#     return(1)
#   }
# }
# ttr_categorizer <- function(data) {
#   if (data >= 103 & data < 190){
#     return("early")
#   } else if (data >= 190 & data < 243){
#     return("middle")
#   } else if (data >= 243 & data <= 521){
#     return("late")
#   }
# }
pdata$`ttr_group` <- sapply(as.numeric(pdata$`Time-to-reccurence (days)`), ttr_categorizer)
ggpaired(data = pdata, id = "Number.x", x = "Manifestation", y = "log2_conc", fill = "Manifestation", palette = colfunc(length(unique(pdata$`ttr_group`))), line.color = "ttr_group", line.size = 0.4) +
  # stat_compare_means(paired = TRUE, label.x = 1.5, geom = "label") +
  # stat_compare_means(comparisons = my_comparisons, group.by="gene", label.y = c(4, 7, 7), label.x = c(1, 2, 3)) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(length(unique(pdata$Manifestation)), "Dark2"), labels = c("Ctrl", "Initial manifestation", "Second manifestation")) +
  # geom_jitter(aes(label=Number.x)) +
  # ggtitle("Initial vs Recurrence") +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="GBM\nOccurence", ncol=1),
         col=guide_legend(title="Time to Recurrence")) -> p1

p1 + 
  stat_pvalue_manual(stat_test, label = "{p.signif} p = {p}", step.group.by = "gene", step.increase = 0.2, hide.ns = T, size=4, bracket.size=1) +
  geom_point(data=data.frame(Manifestation=c("iGBM", "iGBM", "iGBM", "iGBM"), gene=c("ASAH1", "SYNEMIN", "GPNMB", "MMP-9" ), y=c(8, 18, 9, 12)),
             aes(x=Manifestation, y=y, group=gene), alpha=0) -> p1
# p1

## Average Conc per ttr
pdata %>%
  dplyr::filter(Manifestation!="Ctrl") %>%
  dplyr::select(gene, Number.x, Manifestation, log2_conc, ttr_group) %>%
  dplyr::mutate(Manifestation = plyr::mapvalues(Manifestation, from=c("iGBM", "rGBM"), to=c("initial\nGBM", "recurrent\nGBM"))) %>%
  dplyr::group_by(gene, Manifestation, ttr_group) %>%
  dplyr::mutate(avg_log2_conc_per_mani_ttr = median(log2_conc, na.rm = T)) -> pdata_ttr

p1_avg_ttr <- ggplot(data = pdata_ttr, aes(x=Manifestation, y=log2_conc)) +
  geom_boxplot() +
  geom_line(aes(x=Manifestation, y=avg_log2_conc_per_mani_ttr, col=ttr_group, group=ttr_group), inherit.aes = FALSE) +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="GBM\nOccurence", ncol=1),
         col=guide_legend(title="Avg. log2 ( concentration )\nTime to Recurrence"))

@@

## Time to Recurrence ----
train_df_Sub <- pdata %>%
  dplyr::filter( !(Number.x %in% c("32", "33"))) %>%
  dplyr::filter(Manifestation!="Ctrl") %>%
  dplyr::select(Number.x, Time.to.reccurence..days.=`Time-to-reccurence (days)`, class=Manifestation, gene, log2_conc) %>%
  # dplyr::mutate(Time.to.reccurence..days. = ifelse(class=="iGBM", 0, Time.to.reccurence..days.)) %>%
  # dplyr::mutate(class = ifelse(class=="iGBM", 0, 1) ) %>%
  dplyr::rowwise() %>%
  # dplyr::mutate(ttr_group = ttr_categorizer(as.numeric(Time.to.reccurence..days.))) %>%
  # dplyr::mutate(Time.to.reccurence..days. = ttr_categorizer(as.numeric(Time.to.reccurence..days.), 253.5)) %>%
  # dplyr::select(-Time.to.reccurence..days.) %>%
  dplyr::mutate(gene=make.names(gene)) %>%
  tidyr::pivot_wider(id_cols =c("Number.x", "Time.to.reccurence..days.", "class"), names_from = "gene", values_from = "log2_conc") %>%
  # tidyr::pivot_wider(id_cols = c("Number.x", "Time.to.reccurence..days."), names_from = c("class"), values_from = c("ASAH1", "SYNEMIN", "GPNMB", "MMP.9")) %>%
  # dplyr::mutate(ASAH1 = (ASAH1_rGBM+1) /( ASAH1_iGBM+1),
  #               SYNEMIN = (SYNEMIN_rGBM+1)/(SYNEMIN_iGBM+1),
  #               GPNMB = (GPNMB_rGBM+1)/(GPNMB_iGBM+1),
  #               MMP.9 = (MMP.9_rGBM+1)/(MMP.9_iGBM+1)) %>%
  # dplyr::select(Number.x, Time.to.reccurence..days., ASAH1, SYNEMIN, GPNMB, MMP.9 ) %>% 
  # unique() %>%
  dplyr::arrange(Time.to.reccurence..days.)
train_df_Sub <- na.omit(train_df_Sub)
train_df_Sub$Number.x <- as.factor(train_df_Sub$Number.x)

train_df_Sub %>%
  dplyr::group_by(Number.x, Time.to.reccurence..days.) %>%
  dplyr::mutate(ASAH1_delta = ASAH1[class=="rGBM"] - ASAH1[class=="iGBM"],
                SYNEMIN_delta = SYNEMIN[class=="rGBM"] - SYNEMIN[class=="iGBM"],
                GPNMB_delta = GPNMB[class=="rGBM"] - GPNMB[class=="iGBM"],
                MMP.9_delta = MMP.9[class=="rGBM"] - MMP.9[class=="iGBM"]) %>%
  dplyr::rowwise() %>%
  # dplyr::mutate(Time.to.reccurence..days. = ttr_categorizer(as.numeric(Time.to.reccurence..days.))) %>%
  # dplyr::mutate(Time.to.reccurence..days. = ttr_categorizer(as.numeric(Time.to.reccurence..days.), 207.5)) %>%
  dplyr::select(Number.x, Time.to.reccurence..days., ASAH1_delta, SYNEMIN_delta, GPNMB_delta, MMP.9_delta) %>%
  unique() -> ttr_delta

# ttr_delta$Time.to.reccurence..days. <- as.numeric(as.factor(ttr_delta$Time.to.reccurence..days.))
lm_ttr <- lm(Time.to.reccurence..days. ~ ASAH1_delta + SYNEMIN_delta + GPNMB_delta + MMP.9_delta, data=ttr_delta )
summary(lm_ttr)
anova(lm_ttr)

power_calc <- pwr.f2.test(u = 4, f2 = sqrt(summary(lm_ttr)$adj.r.squared), sig.level = 0.05, power = 0.8)
print(sprintf("The estimated sample size needed for 80%% power is: %s", round(power_calc$v) + 5))

ttr_categorizer <- function(data) {
  if (data >= 103 & data < 190){
    return("early")
  } else if (data >= 190 & data < 243){
    return("middle")
  } else if (data >= 243 & data <= 521){
    return("late")
  }
}

train_df_Sub %>%
  dplyr::group_by(Number.x, Time.to.reccurence..days.) %>%
  dplyr::mutate(ASAH1_delta = ASAH1[class=="rGBM"] - ASAH1[class=="iGBM"],
                SYNEMIN_delta = SYNEMIN[class=="rGBM"] - SYNEMIN[class=="iGBM"],
                GPNMB_delta = GPNMB[class=="rGBM"] - GPNMB[class=="iGBM"],
                MMP.9_delta = MMP.9[class=="rGBM"] - MMP.9[class=="iGBM"]) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Time.to.reccurence..days. = ttr_categorizer(as.numeric(Time.to.reccurence..days.))) %>%
  # dplyr::mutate(Time.to.reccurence..days. = ttr_categorizer(as.numeric(Time.to.reccurence..days.), 207.5)) %>%
  dplyr::select(Number.x, Time.to.reccurence..days., ASAH1_delta, SYNEMIN_delta, GPNMB_delta, MMP.9_delta) %>%
  unique() -> ttr_delta

tmp_pdata <- ttr_delta %>% tidyr::pivot_longer(cols = c(ASAH1_delta, SYNEMIN_delta, GPNMB_delta, MMP.9_delta), names_to="gene", values_to="log2_conc")
tmp_pdata$Time.to.reccurence..days. <- factor(tmp_pdata$Time.to.reccurence..days., levels=c("early", "middle", "late"))
ggplot(data = tmp_pdata, aes(x=Time.to.reccurence..days., y=log2_conc, col=Time.to.reccurence..days.)) +
  geom_point() +
  labs(x="", y="log2( Concentration )") +
  facet_wrap(~gene, scales = "free_y") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size=16),
        strip.text = element_text(size=16),
        legend.title = element_text(size=16),
        legend.text = element_text(size=15),
        legend.position = "bottom", legend.box="vertical",) +
  guides(fill=guide_legend(title="GBM\nOccurence", ncol=1),
         col=guide_legend(title="Time to Recurrence"))

@@

# Master figure ----
# subfigure_6 <- ggarrange(p1, p2, p3, ncol = 3)
subfigure_6 <- p1
subfigure_6
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a.png", plot = subfigure_6, width = 21, height = 11, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a.svg", plot = subfigure_6, width = 21, height = 11, dpi = 200)

p1_avg_ttr
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a_avg.png", plot = p1_avg_ttr, width = 21, height = 11, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_a_avg.svg", plot = p1_avg_ttr, width = 21, height = 11, dpi = 200)
```

@@

# train_df_Sub <- train_df %>%
#   dplyr::select(Number.x, Time.to.reccurence..days., class, ASAH1, SYNEMIN, GPNMB, MMP.9) %>%
#   tidyr::pivot_wider(id_cols = c("Number.x", "Time.to.reccurence..days."), names_from = c("class"), values_from = c("ASAH1", "SYNEMIN", "GPNMB", "MMP.9")) %>%
#   dplyr::arrange(Time.to.reccurence..days.)
#   train_df_Sub <- na.omit(train_df_Sub)
# 
# train_df_Sub$Number.x <- as.factor(train_df_Sub$Number.x)
# train_df_Sub$Time.to.reccurence..days. <- as.numeric(train_df_Sub$Time.to.reccurence..days.)
# lm_ttr <- lmerTest::lmer(Time.to.reccurence..days. ~ ASAH1_iGBM + ASAH1_rGBM + SYNEMIN_iGBM + SYNEMIN_rGBM + GPNMB_iGBM + GPNMB_rGBM + MMP.9_iGBM + MMP.9_rGBM + (1 | Type.of.resection), data=train_df_Sub )
# summary(lm_ttr)
# anova(lm_ttr)
# 
# lm_ttr <- lm(Time.to.reccurence..days. ~ ASAH1_iGBM + ASAH1_rGBM + SYNEMIN_iGBM + SYNEMIN_rGBM + GPNMB_iGBM + GPNMB_rGBM + MMP.9_iGBM + MMP.9_rGBM, data=train_df_Sub )
# summary(lm_ttr)
# car::Anova(lm_ttr, type="III")

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_e.png", plot = p_ratio_ttr, width = 15, height = 9, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_e.svg", plot = p_ratio_ttr, width = 15, height = 9, dpi = 200)

@@

## Only use proteins for prediction
data_wide %>%
  mutate(n = row_number()) %>%
  dplyr::select(Number.x, class, n, ASAH1, `MMP-9`, SYNEMIN, GPNMB, `Tumor localization`, Sex, `Age at surgery (years)`, `Type of resection`, `Time-to-reccurence (days)`) -> data_wide
data_wide %>%
  dplyr::mutate(`Time-to-reccurence (days)` = ifelse(class=="iGBM", 0, `Time-to-reccurence (days)`)) -> data_wide

@@

# iGBM vs rGBM ratio  with clinical data ----
# data_long %>% 
#   dplyr::mutate(conc = conc+1) %>% # Add one to avoid cases of 0 division
#   dplyr::group_by(gene, Number.x) %>%
#   dplyr::group_map(function(.d, ...){
#     .d <- as.data.table(.d)
#     .d %>% 
#       dplyr::group_by(gene, Number.x, patient_state, `Age at surgery (years)`, Sex, `Tumor localization`, `Type of resection`, `MGMT promotor methylation`, `p53 accumulation`, `Ki67-Li`, `EGFR vIII`, Treatment, `Time-to-reccurence (days)`, `Latency (days)`) %>% 
#       dplyr::summarise(agg_lab_id = paste(LAB_ID, collapse=";"),
#                        agg_collection_type = paste(collection_type, collapse = ";"), .groups = "keep") -> tmp
#     # print(.d)
#     if ( nrow(.d)==2 & any(!is.na(.d$manifestation_type)) ){
#       .d_initial <- .d[manifestation_type=="Initial manifestation"]
#       .d_second <- .d[manifestation_type=="Second manifestation"]
#       tmp$ratio <- .d_initial$conc / .d_second$conc
#     } else if (nrow(.d)==2) {
#       .d_initial <- .d[collection_type=="E1"]
#       .d_second <- .d[collection_type=="E2"]
#       tmp$ratio <- .d_initial$conc / .d_second$conc
#     } else {
#       tmp$ratio <- .d$conc
#     }
#     return(tmp)
#   }, .keep = TRUE) %>%
#   rbindlist() %>%
#   dplyr::mutate(sample_comparison = plyr::mapvalues(patient_state, from=c(0, 1, 2), to=c("ctrl", "iGBMoverrGBM", "pre-surgery/post-surgery"))) %>%
#   dplyr::filter(sample_comparison=="iGBMoverrGBM" ) %>%
#   dplyr::mutate(log2_conc = log2(ratio)) %>%
#   # dplyr::group_by(gene, Number.x) %>%
#   # dplyr::add_count() %>%
#   # dplyr::filter(n==2) %>%
#   dplyr::mutate(class = plyr::mapvalues(patient_state, from=c(1), to=c("iGBMoverrGBM"))) -> pdata
# 
# data_long %>% 
#   dplyr::mutate(conc = conc+1) %>% # Add one to avoid cases of 0 division
#   dplyr::filter(!is.na(manifestation_type) | patient_state!=0) %>%
#   dplyr::group_by(gene, Number.x) %>%
#   dplyr::group_map(function(.d, ...){
#     .d <- as.data.table(.d)
#     .d %>% 
#       dplyr::group_by(gene, Number.x, patient_state, `Age at surgery (years)`, Sex, `Tumor localization`, `Type of resection`, `MGMT promotor methylation`, `p53 accumulation`, `Ki67-Li`, `EGFR vIII`, Treatment, `Time-to-reccurence (days)`, `Latency (days)`) %>% 
#       dplyr::summarise(agg_lab_id = paste(LAB_ID, collapse=";"),
#                        agg_collection_type = paste(collection_type, collapse = ";"), .groups = "keep") -> tmp
#     # print(.d)
#     if ( nrow(.d)==2 & any(!is.na(.d$manifestation_type)) ){
#       .d_initial <- .d[manifestation_type=="Initial manifestation"]
#       .d_second <- .d[manifestation_type=="Second manifestation"]
#       tmp$ratio <- .d_second$conc / .d_initial$conc
#     } else if (nrow(.d)==2) {
#       .d_initial <- .d[collection_type=="E1"]
#       .d_second <- .d[collection_type=="E2"]
#       tmp$ratio <- .d_initial$conc / .d_second$conc
#     } else {
#       tmp$ratio <- .d$conc
#     }
#     return(tmp)
#   }, .keep = TRUE) %>%
#   rbindlist() %>%
#   dplyr::mutate(sample_comparison = plyr::mapvalues(patient_state, from=c(0, 1, 2), to=c("ctrl", "rGBMoveriGBM", "pre-surgery/post-surgery"))) %>%
#   dplyr::filter(sample_comparison=="rGBMoveriGBM" ) %>%
#   dplyr::filter(grepl(";", agg_lab_id)) %>%
#   dplyr::mutate(log2_conc = log2(ratio)) %>%
#   # dplyr::group_by(gene, Number.x) %>%
#   # dplyr::add_count() %>%
#   # dplyr::filter(n==2) %>%
#   dplyr::mutate(class = plyr::mapvalues(patient_state, from=c(1), to=c("rGBMoveriGBM"))) -> pdata2
# 
# pdata %>%
#   dplyr::select(Number.x, gene, log2_conc) %>%
#   tidyr::pivot_wider(id_cols = "Number.x", names_from = "gene", values_from = "log2_conc") %>%
#   dplyr::mutate(class="iGBM") -> pdata
# 
# pdata2 %>%
#   dplyr::select(Number.x, gene, log2_conc) %>%
#   tidyr::pivot_wider(id_cols = "Number.x", names_from = "gene", values_from = "log2_conc") %>%
#   dplyr::mutate(class="rGBM") -> pdata2
# 
# 
# pdata <- rbindlist(list(pdata, pdata2))
# 
# pdata <- merge(pdata, patient_meta_data[, .(Number.x, `Age at surgery (years)`, Sex, `Tumor localization`, `Type of resection`, `Time-to-reccurence (days)`)], by="Number.x", all.x=T)
# 
# ## Pivot Table
# pdata %>%
#   dplyr::select(Number.x, class, ASAH1, SYNEMIN, GPNMB, `MMP-9`, Sex, `Tumor localization`, Sex, `Age at surgery (years)`, `Type of resection`, `Time-to-reccurence (days)`) -> data_wide
# # data_wide %>%
# #   dplyr::mutate(`Time-to-reccurence (days)` = ifelse(class=="iGBM", 0, `Time-to-reccurence (days)`)) -> data_wide
# 
# ## Prediction
# train_df <- data_wide
# train_df <- na.omit(train_df)  %>% dplyr::select(-Number.x, )
# train_df <- train_df[ !is.infinite(train_df$ASAH1) , ]
# ## Make Valid Column Names 
# colnames(train_df) <- make.names(colnames(train_df))
# train_df$class <- factor(train_df$class, levels=c("iGBM", "rGBM"))
# colnames(train_df) <- make.names(colnames(train_df))
# train_df <- as.data.frame(train_df)
# pred_obj_1.6 <- run_predictions(train_df, pred_col_prob = "rGBM")
# 
# ## ROC Figure 
# g <- ggplot(pred_obj_1.6$roc_data, aes(m=rGBM, d=factor(obs, levels = c("iGBM", "rGBM")), color=Model)) +
#   geom_roc(n.cuts=0) + 
#   scale_color_manual(values = RColorBrewer::brewer.pal(length(unique(pred_obj_1.3$master_lift$liftModelVar)), "Dark2"), labels=c("Logit", "Elastic Logit", "KNN", "RF")) +
#   coord_equal() +
#   style_roc(x="1 - Specificity", y="Sensitivity")
# 
# auc_tbl_1.6 <- as.data.table(calc_auc(g))
# 
# g <- g + annotate("text", x=0.70, y=0.45, label=paste("AUC =", round(auc_tbl_1.6[Model=="Logit"]$AUC, 4)), color=RColorBrewer::brewer.pal(length(unique(pred_obj_1.6$master_lift$liftModelVar)), "Dark2")[1])
# g <- g + annotate("text", x=0.70, y=0.35, label=paste("AUC =", round(auc_tbl_1.6[Model=="Elastic Logit"]$AUC, 4)), color=RColorBrewer::brewer.pal(length(unique(pred_obj_1.6$master_lift$liftModelVar)), "Dark2")[2])
# g <- g + annotate("text", x=0.70, y=0.25, label=paste("AUC =", round(auc_tbl_1.6[Model=="KNN"]$AUC, 4)), color=RColorBrewer::brewer.pal(length(unique(pred_obj_1.6$master_lift$liftModelVar)), "Dark2")[3])
# g <- g + annotate("text", x=0.70, y=0.15, label=paste("AUC =", round(auc_tbl_1.6[Model=="RF"]$AUC, 4)), color=RColorBrewer::brewer.pal(length(unique(pred_obj_1.6$master_lift$liftModelVar)), "Dark2")[4])
# subfigure_1.6 <- g
# 
# pred_obj_1.6$roc_data %>%
#   dplyr::rename(ctrl=iGBM, GBM=rGBM) %>%
#   dplyr::mutate(pred=plyr::mapvalues(pred, from=c("iGBM", "rGBM"), to=c("ctrl", "GBM"))) %>%
#   dplyr::mutate(obs=plyr::mapvalues(obs, from=c("iGBM", "rGBM"), to=c("ctrl", "GBM"))) -> tmp_pred_obj_1.6_roc_data
# 
# ## Merged Plot
# pred_obj_1.6$roc_data %>%
#   dplyr::rename(ctrl=iGBM, GBM=rGBM) %>%
#   dplyr::mutate(pred=plyr::mapvalues(pred, from=c("iGBM", "rGBM"), to=c("ctrl", "GBM"))) %>%
#   dplyr::mutate(obs=plyr::mapvalues(obs, from=c("iGBM", "rGBM"), to=c("ctrl", "GBM"))) -> tmp_pred_obj_1.6_roc_data

@@

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.png", plot = g, width = 11, height = 8.5, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_d.svg", plot = g, width = 11, height = 8.5, dpi = 200)
subfigure_1 <- g
```

@@

<!-- # Latency Data -->
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=9, include=FALSE}
latency_data_file <- "../../data/elisa_plasma_proteomics/ELISA_LATENCY_KM.csv"
latency_data <- fread(latency_data_file)
# Clean names
latency_data <- rstatix::make_clean_names(latency_data)
latency_data %>%
  dplyr::mutate(ASAH1.ng.ml = log2(ASAH1.ng.ml+1),
                SYNM.ng.ml = log2(SYNM.ng.ml+1),
                GPNMB.ng.ml = log2(GPNMB.ng.ml+1),
                MMP.9.ng.ml = log2(MMP.9.ng.ml+1)) %>%
  dplyr::mutate(ASAH1 = ifelse(ASAH1.ng.ml>=mean(ASAH1.ng.ml), "high", "low"),
                SYNM = ifelse(SYNM.ng.ml>=mean(SYNM.ng.ml), "high", "low"),
                GPNMB = ifelse(GPNMB.ng.ml>=mean(GPNMB.ng.ml), "high", "low"),
                MMP9 = ifelse(MMP.9.ng.ml>=mean(MMP.9.ng.ml, na.rm=T), "high", "low")) -> latency_data

fit <- survfit(Surv(Latency..days., status) ~ ASAH1, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p1 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          )

fit <- survfit(Surv(Latency..days., status) ~ SYNM, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p2 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          )

fit <- survfit(Surv(Latency..days., status) ~ GPNMB, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p3 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          )

fit <- survfit(Surv(Latency..days., status) ~ MMP9, data = latency_data)
print(fit)

# Change color, linetype by strata, risk.table color by strata
surv_p4 <- ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = "Dark2"
          )

survival_plot_grid <- ggarrange(surv_p1$plot, surv_p2$plot, surv_p3$plot, surv_p4$plot, ncol = 4)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c.png", plot = survival_plot_grid, width = 18, height = 7, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c.svg", plot = survival_plot_grid, width = 18, height = 7, dpi = 200)

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c1.png", plot = surv_p1$plot, width = 11, height = 7, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c1.svg", plot = surv_p1$plot, width = 11, height = 7, dpi = 200)

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c2.png", plot = surv_p2$plot, width = 11, height = 7, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c2.svg", plot = surv_p2$plot, width = 11, height = 7, dpi = 200)

ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c3.png", plot = surv_p3$plot, width = 11, height = 7, dpi = 200)
ggsave("../../figures/elisa_plasma_proteomics/elisa_main_subfigure_c3.svg", plot = surv_p3$plot, width = 11, height = 7, dpi = 200)
